---
title: "Setup"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
---

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

<br>

## Software installation

Make sure you have installed:

  - **R** -- [link](https://www.r-project.org/)
  - **R studio** -- [link](https://www.rstudio.com/)

----

## R packages installation

Open R studio and run the following lines to install the packages we are going to use during this tutorial.

```{r, eval=FALSE}
install.packages("tidyverse")
install.packages("dada2")
install.packages("phyloseq")
install.packages("DECIPHER")
install.packages("phangorn")
install.packages("ape")
install.packages("seqinr")
```

----

## Organize folders and files

First we need to create a folder named `project` with this structure:

```
project
  |___data
```

Now, we can move and move all our `.fastq.gz` to the folder named `data`, create a `.txt` file with the metadata and a empty `R` file where we will store our code. The project folder will look like this:

```
project
|___data
|     | file1_R1.fastq.gz
|     | file1_R2.fastq.gz
|     | ....... .fastq.gz
|
|___metadata.txt
|___rcode.R
```

<br>

### Step 1. Download the data

Normally, you would just copy the data from your long-term storage device to the folder `project/data`. In this case we will download the data from GitHub. 

[Download this file](https://downgit.github.io/#/home?url=https://github.com/amalacrino/WWU_tutorial/tree/main/project/data) and unzip it inside the folder `project/data`.

<br>

### Step 2. Create the metadata file

Within the project folder, you need to save a file named `metadata.txt` (tab-separated values). The structure of this file should look like this:

```
SampleID    Factor1       Factor2   ...   FactorN
S1          Control       Root      ...
S2          Treatment     Root      ...
```

For this tutorial, you can download one from [here](https://raw.githubusercontent.com/amalacrino/WWU_tutorial/main/project/metadata.txt) and place it inside the `project/` folder. 

<br>

### Step 3. Create a R file

The quickest way to do so is to:

  - Open R Studio
  - Click on `File/New File/ R Script`
  - Click on `File/Save`
  - Navigate to your project folder and save it
  - Close R Studio
  - Open the file you just saved. In this way R Studio will open the file with that relative path already set, so you do not have to worry about changing your code every time you move the folder or when starting a new project.